---
title: "Growth of Japanese quail"
format:
  html:
    toc: true
    html-math-method: katex
    embed-resources: true
    monobackgroundcolor: white
knitr:
  opts_chunk: 
    echo: true
    fig.align: center
highlight-style: github-dark
---


In this example you will analyse a data set of weights of Japanese quail using a hierarchical GAM.

Start by loading the packages that you'll need

```{r}
#| label: packages
#| message: false
pkgs <- c("mgcv", "gratia", "readxl", "scales", "ggplot2", "forcats",
          "dplyr")
vapply(
  pkgs, library, logical(1),
  character.only = TRUE, logical.return = TRUE
)
```

and then load the data. The data were made available as an Excel workbook containing several worksheets. We can use the *readxl* package and its `excel_sheets()` function to look at the contents of the workbook
```{r}
fn <- "quail.xlsx"
excel_sheets(fn)
```
The data we want for this example are in the `"growth analysis"` sheet

```{r}
q_growth <- read_xlsx(fn, sheet = "growth analysis", na = "NA") |>
  mutate(                         # recode some variables as factors
    motherID = factor(motherID),
    eggID = factor(eggID),
    group = factor(group),
    sex = factor(sex)
  ) |>
  # make the label prettier for graphs
  mutate(
    sex = fct_recode(
      sex,
      "Female" = "F",
      "Male" = "M"
    ),
    group = fct_recode(
      group,
      "T[3]" = "T3",
      "T[4]" = "T4",
      "T[3]~T[4]" = "T3T4"
    )
  )
```

::: {.callout-note}

## Task

Fit a factor `by` model to the data in order to model the time-structured treatment mean effects only. We'll ignore the repeated measures aspect of the data for now.

:::